---
title: "fdep_small_streams"
author: "Mike McManus"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Log

2025-06-25 Jay Silvanima at FDEP emailed me Zone1_Staus_FDEP2.txt, which has Florida Albers coordinates for stations. Those columns are GIS_ALBX and GIS_ALBY.

## Libraries

```{r libraries}
library(tidyverse)
library(janitor)
library(sf)
library(mapview)
library(leaflet)
library(leafpop) # for popups in mapview
```

## Read in Data

This website <https://learn.r-journalism.com/en/importing_exporting/delimited_text/delimited-text/> gave example with pipe-delimited text file. What are the coordinates for LONGITUDE and LATITUDE? Are they Florida state plane or UTMs?

```{r fdep_status_data}
# ss_1 <- read_delim(file = "C:/Users/atweg/Documents/Mike/Nokuse/FDEP/Zone1_Status_FD2.txt", delim = "|" )

ss_1 <- read_delim(file = "FDEP/Zone1_Status_FD2.txt", delim = "|" )

names(ss_1)
distinct(ss_1, WATER_RESOURCE)
distinct(ss_1, PARAMETER)
distinct(ss_1, HYDROLOGIC_UNIT_CODE) # n = 18
distinct(ss_1, TMDL_BASIN) # n = 7, TMDL_BASIN > HYDROLOGIC_UNIT_CODE
distinct(ss_1, PLANNING_UNIT) %>% print(n = Inf) # n = 25


```

## Create Simple Features Spatial Data

Site showing switched coordinates <https://stackoverflow.com/questions/61443551/why-sfst-transform-returns-empty-geometries-when-applied-on-a-df-with-sf-colum>.

```{r sf_data}
# my preference to have lowercase column names
ss_2 <- clean_names(ss_1) %>% 
  mutate(.,
               long = gis_albx,
               lat = gis_alby)
names(ss_2)
head(ss_2$long)
head(ss_2$lat)
class(ss_2$long)


#96. Create sf object from csv file with coordinates

#N.B. crs = 3087 is Florida Albers projections
ss_3 <- st_as_sf(ss_2, coords = c( "long", "lat"), crs =3087) 

class(ss_3)
head(ss_3$geometry)

mapview(ss_3)


```

## EDA and Mapview Parameters

```{r eda_maps}

turb_1 <- filter(ss_3, parameter == "Turbidity, Lab")

ggplot(turb_1, aes( y = planning_unit, x = value)) + geom_boxplot() + ggtitle("Turbidity")

#98.  Mapview with bubble and legend pop-up
mapview(turb_1, zcol = "value", cex = "value", alpha.regions = .8, legend = TRUE, popup = popupTable(turb_1, zcol = c("value", "station_name", "waterbody_name", "planning_unit")))

sc_1 <- filter(ss_3, parameter == "Specific Conductance, Field")

ggplot(sc_1, aes( y = planning_unit, x = value)) + geom_boxplot() + ggtitle("Specific Conductivity")

ggplot(sc_1, aes( y = planning_unit, x = log(value))) + geom_boxplot() + ggtitle("Specific Conductivity")

mapview(sc_1, zcol = "value", cex = "value", alpha.regions = .8, legend = TRUE, popup = popupTable(sc_1, zcol = c("value", "station_name", "waterbody_name", "planning_unit")))

ph_1 <- filter(ss_3, parameter == "pH, Field")

ggplot(ph_1, aes( y = planning_unit, x = value)) + geom_boxplot() + ggtitle("pH")

mapview(ph_1, zcol = "value", cex = "value", alpha.regions = .8, legend = TRUE, popup = popupTable(ph_1, zcol = c("value", "station_name", "waterbody_name", "planning_unit")))


sd_1 <- filter(ss_3, parameter == "Transparency (Secchi Depth)")

# filtering b/c value = 19.70 m likely typo as sample_depth is 0.05
ggplot(sd_1 |> filter(value < 19), aes( y = planning_unit, x = value)) + geom_boxplot() + ggtitle("Secchi Depth")

mapview(sd_1, zcol = "value", cex = "value", alpha.regions = .8, legend = TRUE, popup = popupTable(sd_1, zcol = c("value", "station_name", "waterbody_name", "planning_unit")))


```
