---
title: "02_nokuse_sevenruns_adjust_weights"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Adjusting Weights of Nokuse Seven Runs Survey

The purpose of this script is adjust the weights of the Nokuse Seven Runs stream survey that was conducted from September 8-12, 2025. The vignette "Weight Adjustments in Practice" at <https://usepa.github.io/spsurvey/> describes hows to adjust weights for a stratified, equal probability survey, which is the design we used.

We had a sample frame (sr_sf) with a stratum of stream type with four levels: 1)mkc, McKinnon Branch, 2) mst_a, Mainstem above upstream of the Mckinnon Branch confluence, 3) mst_b. Mainstem below the McKinnon Branch confluence, and 4) Otr, other tributaries.

| Stream Type | Stream Length (m) | Base n | Overdraw n |
|:------------|------------------:|-------:|-----------:|
| mck         |             3,326 |      4 |          2 |
| mst_a       |             8,526 |      3 |          2 |
| mst_b       |             7,209 |      3 |          2 |
| otr         |            32,586 |     10 |          2 |

Mainstem: all six base sites were used with 3 being upstream of Mckinnon Branch (sr_mst_a, sr_mst_b, and sr_mst_c) and 3 being downstream of Mckinnon Branch (sr_mst_d, sr_mst_e, and sr_mst_f). No overdraw mainstem sites were used.

Mckinnon Branch: had four base sites and two overdraw sites. The uppermost base site was evaluated in the field by Matt Aresco and I on Thursday, August 28, 2025, and we excluded it because it was logistically inaccessible as it took too long to reach. On Monday, September 8, looking at the aerial photograph, Matt and I reviewed the first Mckinnon overdraw site, which was due west of the sand mine. We decided that it too was logistically inaccessible and excluded it. The second McKinnon Branch overdraw site was reviewed and was deemed a suitable replacement. The four samples collected on McKinnon Branch included the second and third-most upstream sites, which were two base sites called mck_br_a and mck_br_b; the second Mckinnon branch overdraw site, called mck_br_c; the most downstream base site on Mckinnon Branch, called mck_br_d. As we were able to keep the designed sample size, n = 4, for Mckinnon Branch no change in weights is needed.

The otr weights were based on a samples size of n = 10. However, even with 2 overdraw otr sites, only 6 otr sites were sampleable: 1 on Palmetto Branch, 3 on Sugar Mill Branch, 1 on Mill Branch, and 1 on Southern Unnamed Tributary 2. Weights need to be adjusted for otr. Because sampleable otr sites now had an n = 6, we need to adjust the weight.

OTR: Sugar Mill Branch: all three base sites were used and called from most upstream to most downstream: sm_br_a, sm_br_b, and sm_br_c.

OTR: Southern Unnamed Tributary 2: had two base sites that were evaluated, s_un_trib_2a being the uppermost site, and s_un_trib_2b, being the most downstream site (which was close to sr_mst_d). In my orange field notebook, I have the uppermost site, s_un_trib_2a being evaluated as too shallow at \~ 0942. The downstream site s_un_trib_2b was sampleable.

OTR: Northen Unnamed Tributary 3: had two base sites that were evaluated. This is the easternmost tributary, and it crosses Rt 81. The uppermost base site fell on private property so it was excluded from sampling evaluation. Consequently, the first other tributary (otr) overdraw site was used to replace this base site (see Carter Branch below). The lowermost northern unnamed tributary base site (n_un_trib3a) was visited and deemed a nontarget site as it was a pond. Zooming in on the site with the OpenTopoMap base map appears to show that pond.

OTR: Palmetto Branch: one base site fell on Palmetto Branch, with the field visit indicating the site, pa_br_trib_a, met the target population criteria so it was sampled.

OTR: Mill Branch: one base site and one overdraw site fell on Mill Branch. The uppermost Mill Branch site, which was the base site, fell on private property so it was excluded from sampling evaluation. The lowermost Mill Branch was the overdraw site. The overdraw site, call mi_br_a, met the target population criteria was used to replace the uppermost base site on this tributary.

OTR: Grape Head Branch: had one base site, gr_br_a that fell on this tributary and it was evaluated as nontarget as it was too shallow.

OTR: Carter Branch: had one overdraw site, ca_br_a, that fell on this tributary and it was evaluated as nontarget as it was too shallow.

In hindsight, the overdraw of otr sites should have been n = 5 instead of n = 2. A good rule of thumb is that the overdraw sample size should be at least half of the base sample size.

## Library

```{r library}

library(spsurvey)
library(tidyverse)
library(readxl)
library(mapview)
library(tmap)
```

## Log

2025-10-21 adjusted weights b/c of nonresponses and brought in survey water quality data

2025-10-13 modified base and overdraw sites csv file so that column entries of field name of sites and evaluation categories could be added in Excel as sites_stratv4. The field name entries will be used to merge with the survey water quality data.

## Site Evaluation Categories

Read modified base and overdraw sites csv file so that it has field name of sites and evaluation categories.

```{r}
sites_stratv3 <- readRDS("outputs_from_code/sites_stratv3.rds")
class(sites_stratv3)
names(sites_stratv3)
str(sites_stratv3)

write.csv(sites_stratv3, file = "outputs_from_code/sites_stratv3.csv")

sites_stratv4 <- read.csv("outputs_from_code/sites_stratv4.csv")
class(sites_stratv4)
names(sites_stratv4)
str(sites_stratv4)

sites_stratv4 <- dplyr::select(sites_stratv4, c(siteID:comment))

sites_stratv3 <- dplyr::select(sites_stratv3, c(siteID, geometry))


sites_stratv5 <- dplyr::left_join(sites_stratv3, sites_stratv4, by = join_by(siteID))
class(sites_stratv5)
names(sites_stratv5)
str(sites_stratv5)
```

## Adjust Weights

Need to adjust weights because had 1) to use overdraw sites, 2) some sites that were inaccessible, 3) some sites that were not part of the target population, and some sites were not needed. In other words, the sample size of the implemented survey differed from the designed survey.

Get warnings on category estimates that simple random sample estimator is used when n\< 4.

```{r adjwts}

distinct(sites_stratv5, siteuse)
distinct(sites_stratv5, stratum)
distinct(sites_stratv5, eval_cat)

str(sites_stratv5)

dplyr::filter(sites_stratv5, siteuse == "Base") %>%
  group_by(., stratum) %>%
  summarize(sum(wgt))

#code from vignette
# sites_strat$adjwgt <- adjwgt(
#   wgt = sites_strat$wgt,
#   wgtcat = sites_strat$AREA_CAT,
#   framesize = c("small" = 135, "large" = 60)
# )

sites_stratv6 <- filter(sites_stratv5, eval_cat != "not_needed")
# note that observations drop to 24

sites_stratv6$adjwgt <- adjwgt(
  wgt = sites_stratv6$wgt,
  wgtcat = sites_stratv6$stratum,
  framesize = c("mck" = 3326, "mst_a" = 8526, "mst_b" = 7209, "otr" = 32586)
)
# now check adjusted weights 
# needed to use both base and overdraw sites with implemented survey
dplyr::filter(sites_stratv6, siteuse == "Base" | siteuse == "Over") %>%
  group_by(., stratum) %>%
  summarize(sum(adjwgt))

# can simplify code
group_by(sites_stratv6, stratum) %>%
  summarize(sum(adjwgt))

#adjusted weights so that with n = 16 surveyed sites, as opposed to n = 20 designed sites still get sum of weights as designed

# We can use these weights to estimate the proportion of the population that is sampleable or not sampleable due to not being part of the target population, inaccessibility, or a lack of landowner permission. We can do this separately
# for each strata:
est_site_categories <- cat_analysis(
  sites_stratv6,
  vars = "eval_cat",
  stratumID = "stratum",
  weight = "adjwgt"
)

# get warning because local neighborhood variance estimator requires an n = 4 and mst_a and mst_b only have n = 3. With n < 4 independent random sample variance estimator is used
warnprnt()
est_site_categories

# note estimate above is using designed coordinates.

# missing at random (MAR) code from vignette
# Next we adjust for nonresponse, again using adjwgtNR():
# sites_strat$adjwgt_nr <- adjwgtNR(
#   sites_strat$adjwgt,
#   MARClass = sites_strat$AREA_CAT,
#   EvalStatus = sites_strat$EVAL_CAT,
#   TNRClass = c("Endangered", "Inaccessible", "Lack_Permission"),
#   TRClass = "Sampleable"
# )

# Next we adjust now for nonresponse, using adjwgtNR():
sites_stratv6$adjwgt_nr <- adjwgtNR(
  sites_stratv6$adjwgt,
  MARClass = sites_stratv6$stratum,
  EvalStatus = sites_stratv6$eval_cat,
  TNRClass = c("not_target", "inaccessible", "lack_permission"),
  TRClass = "sampleable"
)

# note that only sampleable sites have adjwgt_nr > 0
print(sites_stratv6[, c("siteID", "siteuse", "stratum", "eval_cat",
                      "adjwgt", "adjwgt_nr")], n = 24)

# check again that the adjuste nonresponse weights sum to the sample frame
dplyr::filter(sites_stratv6, eval_cat == "sampleable") %>%
  group_by(., stratum) %>%
  summarize(sum(adjwgt_nr))

# The nonresponse adjusted weights sum to the appropriate population totals.
```

## Compare Surveyed vs Designed Locations

Did Charli record GPS coordinates in the field? Coordinates in spreadsheet overlay exactly coordinates from survey design.

```{r actual_vs_designed_locations}
# obs decreases to n = 16
sites_stratv7 <- dplyr::filter(sites_stratv6, eval_cat == "sampleable")

# field data has mainly chr data so try editing in Excel
dt_1 <- read_excel("survey_wqdata/SR_Data_editsv1.xlsx", col_types = c("text", "text", "date", "date", "numeric", "numeric", "text", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "text", "numeric", "numeric", "numeric", "numeric", "text", "text", "text", "text", "text", "text", "text", "text"))
str(dt_1)
warnings()
# See Long and Lat in decimal degrees

coords <- data.frame(longitude  = dt_1$Longitude, latitude = dt_1$Latitude)
head(coords)

# Coordinate Reference Systems (CRS) for Florida Albers
dt_sf1 <- st_as_sf(dt_1, coords = c("Longitude", "Latitude"), crs =4269) %>% st_transform(3087)
class(dt_sf1)

# on September 23, 2025 believe some of time entries imported incorrectly

head(dt_sf1$Site)
head(sites_stratv7$field_name)

mapview(dt_sf1)

sr_sf <- readRDS("outputs_from_code/srsf.RDS")

mapview(dt_sf1) + mapview(sr_sf, color = "blue")

# did Charli record GPS coordinates in the field
mapview(dt_sf1) + mapview(sites_stratv7, color = "yellow") + mapview(sr_sf, color = "blue")

dt_2 <- left_join(sites_stratv7, dt_1, join_by(field_name == Site))
class(dt_2)
names(dt_2)
head(dt_2)
```

##WQ Estimates

```{r wq_est}
# from analysis vignette
# strat_cont_ests <- cont_analysis(
#   NLA_PNW,
#   siteID = "SITE_ID",
#   vars = "BMMI",
#   weight = "WEIGHT",
#   stratumID = "URBAN"
# )

strat_cont_ests <- cont_analysis(
  dt_2,
  siteID = "siteID",
  vars = "pH",
  weight = "adjwgt_nr",
  stratumID = "stratum"
)

strat_cont_ests$Mean

warnprnt()

names(strat_cont_ests)
strat_cont_ests$CDF
strat_cont_ests$Pct
strat_cont_ests$Mean

```

## Map and Histogram

Code came from script 06 of \Mike\R\_code\nmc-2025-model-ssn2 and updating based on <https://tmap.geocompx.org/charts> to add histogram to map.

```{r}


# Observed sites
### doing legend.show = F for polygons as I will be placing sb_legend on Powerpoint

basin <- readRDS("outputs_from_code/basin.rds")

j_map_vsci_obs <- tm_shape(vahusb_poly) + tm_polygons("subbasin", legend.show = F) +
  tm_shape(DFobs3a) + 
  tm_bubbles(
  size = "vsci",
  legend.size.show = FALSE,
  col = "vsci",
  legend.col.is.portrait = FALSE,
  style = "fixed",
  palette = "viridis",
  title.col = "VSCI at Observed Sites 2001-2018 (n = 199)",
  breaks = seq(0, 90, by = 15),
  sizes.legend=seq(0, 90, by=15),
  legend.hist = FALSE,
) + tm_layout(asp=0,
  legend.width = 2, legend.text.size = 1.5, legend.title.size = 3.0,
  frame = FALSE,
  legend.position = c("center", "TOP"),
  title.position = c('center', 'TOP')
) + tm_scale_bar(position = c("center", "BOTTOM", text.size = 1.0)) + tm_compass(position = c("left", "top")) +
  tm_legend(legend.outside=T, legend.outside.position="top",legend.hist.height = 1.0, legend.hist.width = 1.0, legend.hist.size = 1.5)

tmap_save(j_map_vsci_obs, "figures_nmc_2025/j_map_vsci_obs_v1_20250212.jpeg", width = 7, height = 3.5, dpi = 150)


sr_map_hist_wq <- tm_shape(basin) + 
  tm_polygons() +
  tm_title("Seven Runs Watershed Stream Survey:  pH") + 
  tm_compass(position = c("left", "bottom")) +
  tm_scalebar(position = c("left", "bottom")) +
  tm_shape(sr_sf) + tm_lines() + 
  tm_shape(dt_2) + 
  tm_symbols(fill = "pH", 
             fill.chart = tm_chart_histogram())
  

legend.col.is.portrait = FALSE),
  style = "fixed",
  palette = "viridis",
  title.col = "pH at Observed Sites September 2025 (n = 16)",
  breaks = seq(0, 14, by = 2),
  sizes.legend=seq(0, 14, by = 2),
  legend.hist = FALSE,
) + tm_layout(asp=0,
  legend.width = 2, legend.text.size = 1.5, legend.title.size = 3.0,
  frame = FALSE,
  legend.position = c("center", "TOP"),
  title.position = c('center', 'TOP')
) + tm_scale_bar(position = c("center", "BOTTOM", text.size = 1.0)) + tm_compass(position = c("left", "top")) +
  tm_legend(legend.outside=T, legend.outside.position="top",legend.hist.height = 1.0, legend.hist.width = 1.0, legend.hist.size = 1.5)

tmap_save(j_map_vsci_obs, "figures_nmc_2025/j_map_vsci_obs_v1_20250212.jpeg", width = 7, height = 3.5, dpi = 150)

```
